Topic modeling on EU and the USA research projects

### 2017-06-23 thesis_report/draft_thesis.tex
Spelling

This chapter surveys previous work in topic modeling on scientific data. Topic
modeling algorithms provides a suite of algorithms to discover hidden thematic
structure in large collections of texts and can be adapted to many kinds of
data. These algorithms can be used to summarize, visualize and explore a
collection of documents. The papers listed below are applying a Latend Dirichlet
Allocation model in an interesting approach on their corpus.

@@

The paper that mostly infuenced our work is "Finding Scientific Topics", by
T.Griffiths and M.Steyvers.\cite{griffiths_steyvers04} They investigated
abstract data from articles published in the Proceedings of the National Academy
of Sciences (PNAS) from 1991 to 2001 and compared research topics obtained from
topic modeling with existing categories. In our implementation, we also use 
the abstacts of the projects as documents for the model.

@@

If the prior and the posterior distributions are in the same distribution
family, then the prior and posterior are called conjugate distributions and the
prior is called a conjugate prior for the likelihood funtion.

@@

In LDA, each document is represented by a mixture of the topics, with weight
$\theta_w^2$ for topic $z$ in document $w$. These weight have a Dirichlet prior,
parameterized by a hyper-parameter $\alpha$. Each topic is a probability
distribution over the vocabulary, with probability $\beta^w_z$ for word $w$ in
topic $z$. We define the following variables and notation:
\begin{itemize}
  \item[] $k$ is the number of topics.
  \item[] $V$ is the number of unique words in the vocabulary.
  \item[] $\theta$  is the topic distribution (of length $k$) for a document,
   drawn from a uniform Dirichlet distribution with parameter $\alpha$.
  \item[] $z_{n}$ is a topic assignment for $word w_{n}$, sampled from 
  $p(z_{n} = i|\theta) = \theta_{i}$.
  \item[] \textbf{w}$ = (w_{1}, ... , w_{N})$ is a document with $N$ words.
  \item[] $w_{n}^{i} = 1$ means that the word $w_{n}$ is the $i$-th word 
  of the vocabulary.
  \item[] $\beta$  is a $k \times V$ matrix, where each row $\beta_{i}$ 
  is the multinomial distribution for the $i$-th topic. That is, 
  $\beta_{ij} = p(w^{j} = 1 | z_{j} = i)$.
\end{itemize}

@@

As seen in step 1, each document contains topics in different proportions,
determined by the scaling parameter $\alpha$. The individual words are drawn
from one of the $k$ topics in step (b), where the probability of each word
within a topic is parameterized by a $k \times V$ matrix , where $V$ is the
size of the specified vocabulary. The most probable words in each topic could
be used to identify the topic and gives an intuitive description for the topic.

@@

Specifically, we have:
\begin{equation}
q(z_{di}=k) = \phi_{dw_{di}k}
\end{equation}
\begin{equation}
q(\theta_{d}) = dirichlet(\theta_{d}; \gamma_{d})
\end{equation}
\begin{equation}
q(\beta_{k}) = dirichlet(\beta_{k}; \lambda_{k})
\end{equation}

The posterior over the per-word topic assignments $z$ is parameterized by
$\phi$, the posterior over per-document topic weights $\theta$ is parameterized
by $\gamma$, and the posterior over the topics $\beta$ is parameterized by
$\lambda$.

@@

CORDIS (Community Research and Development Information Service) is the European
Commission's primary public repository and portal to disseminate information on
all EU-funded research projects and their results in the broadest sense. The
website includes all public information held by the Commission (project
factsheets, publishable reports and deliverables), editorial content to support
communication and exploitation (news, events, success stories, magazines) and
comprehensive links to external sources such as open access publications and
websites. CORDIS is managed by the Publications Office of the European Union, on
behalf of the European Commission's research Directorates-General and Agencies.
CORDIS content dates back to the origin of the service in 1990 and the website
has been online since 1994.

Below, we provide the general topics of the EU projects as listed in the EU 
Open Data Portal \footnote{https://data.europa.eu/euodp/en/data}:
\begin{itemize}
\item Agriculture, Fisheries and Food
	\begin{description}
	\item Agriculture, Food safety, Maritime affairs and fisheries
	\end{description}
\item Business
	\begin{description}
	\item Competition, Enterprise, Single market, Trade
	\end{description}
\item Culture and Education
	\begin{description}
	\item Audiovisual and media, Culture, Education training and youth, Multiligualism
	\end{description}
\item Customs and Tax
	\begin{description}
	\item Customs, Taxation
	\end{description}
\item Development and Humanitarian aid
	\begin{description}
	\item Development and cooperation, Humanitarian aid and Civil protection, Human rights
	\end{description}
\item Economy and Finance
	\begin{description}
	\item Budget, Economic and monetary affairs, Fraud prevention
	\end{description}
\item Employment and Social affairs
\item Enlargement and Foreign affairs
\item Environment and Energy
	\begin{description}
	\item Climate action, Energy, Environment
	\end{description}
\item EU institutions
	\begin{description}
	\item Institutional affairs
	\end{description}
\item Health
	\begin{description}
	\item Health, Sport
	\end{description}
\item Justice and Citizens’ rights
	\begin{description}
	\item EU citizenship, Consumers, Justice and Home affairs
	\end{description}
\item Regions and Local development
	\begin{description}
	\item Regional policy
	\end{description}
\item Science and Technology
	\begin{description}
	\item Space, Digital Economy and Society, Research and Innovation
	\end{description}
\item Transport and Travel
\end{itemize}

The data that are provided by CORDIS (via the EU Open Data Portal) are open,
free and can be used in research, applications, commercial or non-commercial
purposes. The dataset that we chose for the following analysis contains
information about all the projects funded by the European Union from 1994 to
2017. These projects got approved under the framework programme (FP4-H2020) for
research and technological development. For each project it is provided
information, such as: reference, acronym, dates, funding, programmes,
participant countries, subjects and objectives.

@@

The LDA model requires some basic input parameters; for both datasets we chose
the number of topics $K=10$. The hyperparameter $\alpha$, which affects the
sparsity of the document-topic distribution $\theta$, is by default a symmetric
$\frac{1.0}{num_topics} = 0.1$ prior. We left the default value of $chunksize=2000$. We
started with 50 iterations and later we increased the number of iterations,
until the training converged. We used 7000 iterations for each framework
programme of the EU dataset and 8000 for each framework of the USA dataset. In
the USA dataset, we used $passes=1$, while in the EU dataset we switched to
$passes=2$, because the training could not converge, despite the large number of
the iterations.

@@

In this section, we show our main results. We focus on finding differents
between the two datasets in terms of topic distribution and later we 
investigate the evolution of the topics over time of each dataset separately.

@@

In general, the most popular subject in European Union over the years is
Information systems. In our dataset we fount projects related to almost all
kinds of information systems, and especially Quality managemest systems
(1994-2002), Management information systems, Network information systems and
Information systems as a service to all kinds of organizations, including
universities and companies. Since technology has an impact in every aspect
nowadays, it is normal to have such polularity over the years and it's becoming
popular in education, as well. The same period of time, Innovation in Business
emerged as a trending topic, with projects related to new technologies and
software systems in the industry and how they assist the businesses and the
market. In addition, until 2002 a very popular subject related to Information
systems, was Applications in mobile networks, embedded and broadband systems,
which nowadays has lost part of its popularity in research.

Another popular subject that appears the last 22 years is Materials research;
the projects related to material research are mostly related to the discovery
and design of new materials, with an emphasis on solids. Matelials properties
and performace are used in many disciplines, like chemistry, physics and
engineering.

The third most polular subject is Molecular biology and Genetics. These projects
are related to the cellular molecules that carry out the biological processes
essential for the cell's functions and maintenance. These affect many aspects of
our lives, including evolution, nutrition and the spread of diseases. Especially
from 1998 up to day, there is a focus in Healthcare and Diagnostics, and fot the
last decade researchers are studying the cancer detection and treatment, which
is a common cause of death in EU and the second most common cause of death in
USA.

Moreover, Environmental, Social and Economic Sustainability is a subject very
important to the research community and many researchers work on this up to day.
In short, sustainability takes into account how we might live in harmony with
the natural world around us, protecting it from damage and destruction. Many
projects are related to one of the Three Pillars of Sustainability\footnote{http
://www.thwink.org/sustain/glossary/ThreePillarsOfSustainability.htm}. The
European Union is considered by some to have the most extensive environmental
laws of any international organisation, started by the  European Economic
Community (EEC), which is the beginning of the EU's environmental policy. Social
and Econimic policies are also popular in the EU region over the last decate.
Also, the last two decades there is a focus on the Green energy production, like
hydropower, solar and nuclear energy.

@@

The National Science Foundation mostly supports fundamental research and
education in the non-medicalin science and engineering. Its medical counterpart
is the National Institutes of Health.\footnote{\url{https://en.wikipedia.org/wik
i/National_Science_Foundation}} With STEM being the top priority in USA, most of
the projects founded by NSF are related to engineering, computer science,
mathematics and information systems, in general. The subjects STEM \& Education,
Software engineering, Information systems and Mathematics, arise as topics in
all these years, from 1994-2017.

While NSF promotes the progress of science, it also enhances the effords to
maintain the national health, prosperity, and welfare. Over the last 22 years,
many projects are related to the environment effects of climate change. Topics
like Environmental Science, Atmospheric \& Ocean sciences, Geology, Climate
change and its impact in the United States emerge during all these years. The
last three years there is a focus on the Green energy production, like
hydropower and wind power. Among with climate change, there is many research
related to Biology and Evolution. A less popular subject that appears during the
years 1994-1997 and 2007-2013 is Molecular, Cellular \& Organismal biology and
Genetics. These projects mostly refer to applied genetics, as well as cell
functionality. Topics include receptor biology, regulation of gene expression,
biomedical engineering, patterns of inheritance and molecular genetics.

Another popular subject that appears the last 22 years is Materials and
Chemistry; the projects related to material research are mostly refer to the
materials' properties and applications in engineering. The last 3 years,
researchers from the USA focus on materials applications and desing in
mechanical engineering, bioengineering and chemical \& transport systems. In
2002-2006 many projects are related to the subject of transport that is
sustainable in the senses of social, environmental and climate impacts (Green
transportation).

@@

One of our goals is to identify trends in scientific research in EU and USA, in
order to give some insights to researchers and scientists regarding potential
trends. Fot this analysis, we run the model for the whole dataset, instead of
running the model per FP), for the EU and the USA, respectively. The outcome was
10 topics for each region from 1994 to 2017 for the EU and from 1994 to 2016 for
the USA.\footnote{in this implementation we removed the projects of the year
2017 from the USA dataset, because the amount of projects-documents was too
small compared to the other years and it was affecting the result of the trend
for eact topic.}

@@

In figure x we show the 10 topics that consists of the 20 top words with the
highest posterior probability of belonging to the topic. The size of each word
is in proportion to each probability; the bigger the word, the higher the
probability. Each wordcloud describes a topic, which we can find intuitevely.
Below are listed the topics that we found corresponding to the words that
describe them:
\begin{itemize}
\item[] Topic 0: Molecular biology
\item[] Topic 1: Marerials \& Manufacturing
\item[] Topic 2: Social policy
\item[] Topic 3: Disease detection \& treatment
\item[] Topic 4: Materials science / Physics
\item[] Topic 5: Transport \& Energy
\item[] Topic 6: Business \& Innovation
\item[] Topic 7: Climate change / Environment
\item[] Topic 8: Molecular mechanisms
\item[] Topic 9: Information systems
\item[] Topic 10: International research
\item[] Topic 11: Energy production
\end{itemize}
\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Topic 0]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic0_all.png}} &
\subfloat[Topic 1]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic1_all.png}} &
\subfloat[Topic 2]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic2_all.png}}\\
\subfloat[Topic 3]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic3_all.png}} &
\subfloat[Topic 4]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic4_all.png}} &
\subfloat[Topic 5]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic5_all.png}}\\
\subfloat[Topic 6]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic6_all.png}} &
\subfloat[Topic 7]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic7_all.png}} &
\subfloat[Topic 8]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic8_all.png}}\\
\subfloat[Topic 9]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic9_all.png}} &
\subfloat[Topic 10]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic10_all.png}} &
\subfloat[Topic 11]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic11_all.png}}\\
\end{tabular}
\caption{Wordclouds of topics of the EU dataset (1994-2017)}
\end{figure}
\end{center}
In order to find the topic distribution over time, we need the topic
distribution of each project-document. As we mentioned before, LDA assumes that
each document exhibits multiple topics. After running the model to our dataset,
we get the document-topic distribution and we keep the topic with the highest
probability (the one that most likely belongs to this document). Then, we
calculate the proportion of each topic per FP, using the following equation:
\begin{equation}
topicProportion(k, fp) = \frac{\sum num of docs_{k,fp}}{\sum num of docs_{fp}} * 100
\end{equation}
where $k$ is the number of the topic id, $fp$ is the framework programme,
$fp \in [FP4, FP5, FP6, FP7, H2020]$. The topic proportion of each topic in 
the FP equals to the sum of the number of the documents assigned to this 
topic during the FP, nomralized by the number of the documents-projects 
occuring in this FP. In figure X we plot the results of the topic proportion 
over time.

@@

where $k$ is the number of the topic id, $y$ is the year between 1994-2016. The
topic proportion of each topic per year equals to the sum of the number of the
documents assigned to this topic during the year, nomralized by the number of
the documents-projects  occuring in this year.

In figure X we plot the topic distribition over time. We can easily observe that
the topic 6, which refers to STEM education, is becoming a trend in the United
States over the years, while most of the other topics' evolution doesn't show up
such a high increace. Topics 5 and 8, which refer to Chemistry \& Materials
science and Molecular biology, respectively, show a small decreace over the
years. Topics 9 and 0 (Mechanical Engineering, Materiac Science and High energy
physics) show a small increase over the last 5 years. Mathematics' evolution
over time (Topic 1) looks stable on average and shows up as a popular subject
over the last 22 years, while Social Sciences and International Research (Topic
3 and 4) show a small decrease over the last 3 years.  Molecular biology and
Genetics (Topic 2) has an increasing polularity from 1994 to 2001 and later is
starting to stabilize. Climatology, Atmospheric and Ocean sciences look stable
the first years, while from 2006 up to date it is gradually becoming less
popular over the years.

@@

\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Management software]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic0_7000.png}} &
\subfloat[Hydropower / Energy production]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic1_7000.png}}&
\subfloat[Information technologies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic2_7000.png}}\\
\subfloat[Quality control systems]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic3_7000.png}}&
\subfloat[Healthcare / Medical genetics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic4_7000.png}} &
\subfloat[Environmental sustainability and policies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic5_7000.png}}\\
\subfloat[Molecular biology (genes)]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic6_7000.png}} &
\subfloat[Optical quantum computing]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic7_7000.png}}&
\subfloat[Education in Nuclear physics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic8_7000.png}}\\
\subfloat[Education, training]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP5 (1998-2002)}
\end{figure}
\end{center}

>>>>

This chapter surveys previous work in topic modeling on scientific data. Topic
modeling algorithms provides a suite of algorithms to discover hidden thematic
structure in large collections of texts and can be adapted to many kinds of
data. These algorithms can be used to summarize, visualize and explore a
collection of documents. The papers listed below are applying a Latent Dirichlet
Allocation model in an interesting approach on their corpus.

@@

The paper that mostly influenced our work is "Finding Scientific Topics", by
T.Griffiths and M.Steyvers.\cite{griffiths_steyvers04} They investigated
abstract data from articles published in the Proceedings of the National Academy
of Sciences (PNAS) from 1991 to 2001 and compared research topics obtained from
topic modeling with existing categories. In our implementation, we also use 
the abstracts of the projects as documents for the model.

@@

If the prior and the posterior distributions are in the same distribution
family, then the prior and posterior are called conjugate distributions and the
prior is called a conjugate prior for the likelihood function.

@@

In LDA, each document is represented by a mixture of the topics, with weight
$\theta_w^2$ for topic $z$ in document $w$. These weight have a Dirichlet prior,
parametrized by a hyper-parameter $\alpha$. Each topic is a probability
distribution over the vocabulary, with probability $\beta^w_z$ for word $w$ in
topic $z$. We define the following variables and notation:
\begin{itemize}
  \item[] $k$ is the number of topics.
  \item[] $V$ is the number of unique words in the vocabulary.
  \item[] $\theta$  is the topic distribution (of length $k$) for a document,
   drawn from a uniform Dirichlet distribution with parameter $\alpha$.
  \item[] $z_{n}$ is a topic assignment for $word w_{n}$, sampled from 
  $p(z_{n} = i|\theta) = \theta_{i}$.
  \item[] \textbf{w}$ = (w_{1}, ... , w_{N})$ is a document with $N$ words.
  \item[] $w_{n}^{i} = 1$ means that the word $w_{n}$ is the $i$-th word 
  of the vocabulary.
  \item[] $\beta$  is a $k \times V$ matrix, where each row $\beta_{i}$ 
  is the multinomial distribution for the $i$-th topic. That is, 
  $\beta_{ij} = p(w^{j} = 1 | z_{j} = i)$.
\end{itemize}

@@

As seen in step 1, each document contains topics in different proportions,
determined by the scaling parameter $\alpha$. The individual words are drawn
from one of the $k$ topics in step (b), where the probability of each word
within a topic is parametrized by a $k \times V$ matrix , where $V$ is the
size of the specified vocabulary. The most probable words in each topic could
be used to identify the topic and gives an intuitive description for the topic.

@@

Specifically, we have:
\begin{equation}
q(z_{di}=k) = \phi_{dw_{di}k}
\end{equation}
\begin{equation}
q(\theta_{d}) = Dirichlet(\theta_{d}; \gamma_{d})
\end{equation}
\begin{equation}
q(\beta_{k}) = Dirichlet(\beta_{k}; \lambda_{k})
\end{equation}

The posterior over the per-word topic assignments $z$ is parametrized by
$\phi$, the posterior over per-document topic weights $\theta$ is parametrized
by $\gamma$, and the posterior over the topics $\beta$ is parametrized by
$\lambda$.

@@

CORDIS (Community Research and Development Information Service) is the European
Commission's primary public repository and portal to disseminate information on
all EU-funded research projects and their results in the broadest sense. The
website includes all public information held by the Commission (project
fact sheets, publishable reports and deliverables), editorial content to support
communication and exploitation (news, events, success stories, magazines) and
comprehensive links to external sources such as open access publications and
websites. CORDIS is managed by the Publications Office of the European Union, on
behalf of the European Commission's research Directorates-General and Agencies.
CORDIS content dates back to the origin of the service in 1990 and the website
has been online since 1994.

Below, we provide the general topics of the EU projects as listed in the EU 
Open Data Portal \footnote{https://data.europa.eu/euodp/en/data}:
\begin{itemize}
\item Agriculture, Fisheries and Food
	\begin{description}
	\item Agriculture, Food safety, Maritime affairs and fisheries
	\end{description}
\item Business
	\begin{description}
	\item Competition, Enterprise, Single market, Trade
	\end{description}
\item Culture and Education
	\begin{description}
	\item Audiovisual and media, Culture, Education training and youth, Multilingualism
	\end{description}
\item Customs and Tax
	\begin{description}
	\item Customs, Taxation
	\end{description}
\item Development and Humanitarian aid
	\begin{description}
	\item Development and cooperation, Humanitarian aid and Civil protection, Human rights
	\end{description}
\item Economy and Finance
	\begin{description}
	\item Budget, Economic and monetary affairs, Fraud prevention
	\end{description}
\item Employment and Social affairs
\item Enlargement and Foreign affairs
\item Environment and Energy
	\begin{description}
	\item Climate action, Energy, Environment
	\end{description}
\item EU institutions
	\begin{description}
	\item Institutional affairs
	\end{description}
\item Health
	\begin{description}
	\item Health, Sport
	\end{description}
\item Justice and Citizens’ rights
	\begin{description}
	\item EU citizenship, Consumers, Justice and Home affairs
	\end{description}
\item Regions and Local development
	\begin{description}
	\item Regional policy
	\end{description}
\item Science and Technology
	\begin{description}
	\item Space, Digital Economy and Society, Research and Innovation
	\end{description}
\item Transport and Travel
\end{itemize}

The data that are provided by CORDIS (via the EU Open Data Portal) are open,
free and can be used in research, applications, commercial or non-commercial
purposes. The dataset that we chose for the following analysis contains
information about all the projects funded by the European Union from 1994 to
2017. These projects got approved under the framework programme (FP4-H2020) for
research and technological development. For each project it is provided
information, such as: reference, acronym, dates, funding, programs,
participant countries, subjects and objectives.

@@

The LDA model requires some basic input parameters; for both datasets we chose
the number of topics $K=10$. The hyper-parameter $\alpha$, which affects the
sparsity of the document-topic distribution $\theta$, is by default a symmetric
$\frac{1.0}{num_topics} = 0.1$ prior. We left the default value of $chunksize=2000$. We
started with 50 iterations and later we increased the number of iterations,
until the training converged. We used 7000 iterations for each framework
programme of the EU dataset and 8000 for each framework of the USA dataset. In
the USA dataset, we used $passes=1$, while in the EU dataset we switched to
$passes=2$, because the training could not converge, despite the large number of
the iterations.

@@

In this section, we show our main results. We focus on finding differences
between the two datasets in terms of topic distribution and later we 
investigate the evolution of the topics over time of each dataset separately.

@@

In general, the most popular subject in European Union over the years is
Information systems. In our dataset we fount projects related to almost all
kinds of information systems, and especially Quality management systems
(1994-2002), Management information systems, Network information systems and
Information systems as a service to all kinds of organizations, including
universities and companies. Since technology has an impact in every aspect
nowadays, it is normal to have such popularity over the years and it's becoming
popular in education, as well. The same period of time, Innovation in Business
emerged as a trending topic, with projects related to new technologies and
software systems in the industry and how they assist the businesses and the
market. In addition, until 2002 a very popular subject related to Information
systems, was Applications in mobile networks, embedded and broadband systems,
which nowadays has lost part of its popularity in research.

Another popular subject that appears the last 22 years is Materials research;
the projects related to material research are mostly related to the discovery
and design of new materials, with an emphasis on solids. Materials properties
and performance are used in many disciplines, like chemistry, physics and
engineering.

The third most popular subject is Molecular biology and Genetics. These projects
are related to the cellular molecules that carry out the biological processes
essential for the cell's functions and maintenance. These affect many aspects of
our lives, including evolution, nutrition and the spread of diseases. Especially
from 1998 up to day, there is a focus in Health care and Diagnostics, and for the
last decade researchers are studying the cancer detection and treatment, which
is a common cause of death in EU and the second most common cause of death in
USA.

Moreover, Environmental, Social and Economic Sustainability is a subject very
important to the research community and many researchers work on this up to day.
In short, sustainability takes into account how we might live in harmony with
the natural world around us, protecting it from damage and destruction. Many
projects are related to one of the Three Pillars of Sustainability\footnote{http
://www.thwink.org/sustain/glossary/ThreePillarsOfSustainability.htm}. The
European Union is considered by some to have the most extensive environmental
laws of any international organization, started by the  European Economic
Community (EEC), which is the beginning of the EU's environmental policy. Social
and Economic policies are also popular in the EU region over the last decade.
Also, the last two decades there is a focus on the Green energy production, like
hydropower, solar and nuclear energy.

@@

The National Science Foundation mostly supports fundamental research and
education in the non-medical science and engineering. Its medical counterpart
is the National Institutes of Health.\footnote{\url{https://en.wikipedia.org/wik
i/National_Science_Foundation}} With STEM being the top priority in USA, most of
the projects founded by NSF are related to engineering, computer science,
mathematics and information systems, in general. The subjects STEM \& Education,
Software engineering, Information systems and Mathematics, arise as topics in
all these years, from 1994-2017.

While NSF promotes the progress of science, it also enhances the efforts to
maintain the national health, prosperity, and welfare. Over the last 22 years,
many projects are related to the environment effects of climate change. Topics
like Environmental Science, Atmospheric \& Ocean sciences, Geology, Climate
change and its impact in the United States emerge during all these years. The
last three years there is a focus on the Green energy production, like
hydropower and wind power. Among with climate change, there is many research
related to Biology and Evolution. A less popular subject that appears during the
years 1994-1997 and 2007-2013 is Molecular, Cellular \& Organismal biology and
Genetics. These projects mostly refer to applied genetics, as well as cell
functionality. Topics include receptor biology, regulation of gene expression,
biomedical engineering, patterns of inheritance and molecular genetics.

Another popular subject that appears the last 22 years is Materials and
Chemistry; the projects related to material research are mostly refer to the
materials' properties and applications in engineering. The last 3 years,
researchers from the USA focus on materials applications and design in
mechanical engineering, bioengineering and chemical \& transport systems. In
2002-2006 many projects are related to the subject of transport that is
sustainable in the senses of social, environmental and climate impacts (Green
transportation).

@@

One of our goals is to identify trends in scientific research in EU and USA, in
order to give some insights to researchers and scientists regarding potential
trends. For this analysis, we run the model for the whole dataset, instead of
running the model per FP), for the EU and the USA, respectively. The outcome was
10 topics for each region from 1994 to 2017 for the EU and from 1994 to 2016 for
the USA.\footnote{in this implementation we removed the projects of the year
2017 from the USA dataset, because the amount of projects-documents was too
small compared to the other years and it was affecting the result of the trend
for each topic.}

@@

In figure x we show the 10 topics that consists of the 20 top words with the
highest posterior probability of belonging to the topic. The size of each word
is in proportion to each probability; the bigger the word, the higher the
probability. Each wordcloud describes a topic, which we can find intuitively.
Below are listed the topics that we found corresponding to the words that
describe them:
\begin{itemize}
\item[] Topic 0: Molecular biology
\item[] Topic 1: Materials \& Manufacturing
\item[] Topic 2: Social policy
\item[] Topic 3: Disease detection \& treatment
\item[] Topic 4: Materials science / Physics
\item[] Topic 5: Transport \& Energy
\item[] Topic 6: Business \& Innovation
\item[] Topic 7: Climate change / Environment
\item[] Topic 8: Molecular mechanisms
\item[] Topic 9: Information systems
\item[] Topic 10: International research
\item[] Topic 11: Energy production
\end{itemize}
\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Topic 0]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic0_all.png}} &
\subfloat[Topic 1]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic1_all.png}} &
\subfloat[Topic 2]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic2_all.png}}\\
\subfloat[Topic 3]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic3_all.png}} &
\subfloat[Topic 4]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic4_all.png}} &
\subfloat[Topic 5]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic5_all.png}}\\
\subfloat[Topic 6]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic6_all.png}} &
\subfloat[Topic 7]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic7_all.png}} &
\subfloat[Topic 8]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic8_all.png}}\\
\subfloat[Topic 9]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic9_all.png}} &
\subfloat[Topic 10]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic10_all.png}} &
\subfloat[Topic 11]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic11_all.png}}\\
\end{tabular}
\caption{Wordclouds of topics of the EU dataset (1994-2017)}
\end{figure}
\end{center}
In order to find the topic distribution over time, we need the topic
distribution of each project-document. As we mentioned before, LDA assumes that
each document exhibits multiple topics. After running the model to our dataset,
we get the document-topic distribution and we keep the topic with the highest
probability (the one that most likely belongs to this document). Then, we
calculate the proportion of each topic per FP, using the following equation:
\begin{equation}
topicProportion(k, fp) = \frac{\sum num of docs_{k,fp}}{\sum num of docs_{fp}} * 100
\end{equation}
where $k$ is the number of the topic id, $fp$ is the framework programme,
$fp \in [FP4, FP5, FP6, FP7, H2020]$. The topic proportion of each topic in 
the FP equals to the sum of the number of the documents assigned to this 
topic during the FP, normalized by the number of the documents-projects 
occurring in this FP. In figure X we plot the results of the topic proportion 
over time.

@@

where $k$ is the number of the topic id, $y$ is the year between 1994-2016. The
topic proportion of each topic per year equals to the sum of the number of the
documents assigned to this topic during the year, normalized by the number of
the documents-projects  occurring in this year.

In figure X we plot the topic distribution over time. We can easily observe that
the topic 6, which refers to STEM education, is becoming a trend in the United
States over the years, while most of the other topics' evolution doesn't show up
such a high increase. Topics 5 and 8, which refer to Chemistry \& Materials
science and Molecular biology, respectively, show a small decrease over the
years. Topics 9 and 0 (Mechanical Engineering, Material Science and High energy
physics) show a small increase over the last 5 years. Mathematics' evolution
over time (Topic 1) looks stable on average and shows up as a popular subject
over the last 22 years, while Social Sciences and International Research (Topic
3 and 4) show a small decrease over the last 3 years.  Molecular biology and
Genetics (Topic 2) has an increasing popularity from 1994 to 2001 and later is
starting to stabilize. Climatology, Atmospheric and Ocean sciences look stable
the first years, while from 2006 up to date it is gradually becoming less
popular over the years.

@@

\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Management software]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic0_7000.png}} &
\subfloat[Hydropower / Energy production]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic1_7000.png}}&
\subfloat[Information technologies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic2_7000.png}}\\
\subfloat[Quality control systems]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic3_7000.png}}&
\subfloat[Health care / Medical genetics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic4_7000.png}} &
\subfloat[Environmental sustainability and policies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic5_7000.png}}\\
\subfloat[Molecular biology (genes)]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic6_7000.png}} &
\subfloat[Optical quantum computing]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic7_7000.png}}&
\subfloat[Education in Nuclear physics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic8_7000.png}}\\
\subfloat[Education, training]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP5 (1998-2002)}
\end{figure}
\end{center}
